---
title: "covid 19"
author: "Abel Jacobo Perez"
date: "5/2/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$:dying is independent of sex.
$H_A$: dying is dependent on sex.




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result, Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result, y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result, y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
 
 
  Among those who died in Korea from Covid, most of them were Men. Woman died at almost half the rate as men dead in Korea from Covid. The right bar graph(red) shows woman died at a levels of 36% of the infected,and it shows that that the men(blue) died at levels of 64% of those infected.


  





  
## Numerical results

```{r}
table2 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```
Men died at almost twice the rate as woman did from Covid-19 in Korea. 36% of the dead count were females and 64% of the death count were men.

Very few people actually died , 1.4% of Korean females died however 3% of Korean men died,almost twice the rate as woman.


## Inferential results 

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The p value 0.001 is to low thus we reject the Null hypothesis. The death rate on Korea is dependent on sex.



## Conclusions

